---
title: "Spatial Analysis with leaflet"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "May 19, 2022"
output:
  html_document:
    fig_height: 5
    fig_width: 7
  pdf_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE}
library(tidyverse)   # Load additional packages here 
library(leaflet)
library(macleish)

# Some customization.  You can alter or delete as desired (if you know what you are doing).
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```



### Mapping Amherst

Let's start with a simple example where we create an interactive map at a specified point.

```{r eval = TRUE}
library(leaflet)
m <- leaflet() %>%
  addTiles() %>% # Add OpenStreetMap tiles
  addMarkers(lng = -72.516, lat = 42.369, popup = "Amherst College<br>Founded 1821")
m
```

### Mapping MacLeish

The MacLeish Field Station is a 250 acre patchwork of forest and pasture land in Whately, MA owned and operated by Smith College.
A data package in R is available with information about the station and the surrounding area.


```{r eval = TRUE}
library(macleish)
names(macleish_layers)
summary(macleish_layers[["buildings"]])
```

```{r eval = TRUE}
m <- leaflet() %>%
  addTiles() %>%
  addPolygons(data = macleish_layers[["boundary"]],
              weight = 1) %>%
  addPolygons(data = macleish_layers[["buildings"]],
              weight = 1) %>%
  addMarkers(data = subset(macleish_layers[["landmarks"]], grepl("Met", Label)), 
             popup = ~ Label)
m
```

### Challenge 1:

In groups of two, create a map of the MacLeish field station that includes some other attributes not already used here.  

```{r}
# your solution here...

```


### Challenge 2: 

pdxTrees is a data package composed of information on inventoried trees in Portland, OR.
There are two datasets that can be accessed with this package:

- `get_pdxTrees_parks()` pulls in data on up to 25,534 trees from 174 Portland parks.
- `get_pdxTrees_streets()` pulls in data on up to 218,602 trees located on Portland’s streets. A street tree is loosely defined as a tree generally in the public right-of-way, usually between the sidewalk and the street.

Use the https://github.com/mcconvil/pdxTrees package and leaflet to create an interactive map of trees in public parks in Portland, OR. (See https://teachdatascience.com/datapackage/ for some examples of how to access the data and the RStudio leaflet pages for examples.)  Your map should utilize some feature of leaflet beyond what is included in this lab.  Think carefully about how to include multiple dimensions of data in your interactive map.

```{r}
#remotes::install_github("mcconvil/pdxTrees")
library(pdxTrees)
pdxTrees_parks <- get_pdxTrees_parks()
glimpse(pdxTrees_parks)
```

SUBMISSION INSTRUCTIONS: to receive full credit for this lab activity, please upload both your Rmd file and your formatted html file to Moodle (one submission per group).
Both of your names must be included in the yaml header.
Deadline is midnight on Tuesday, November 12th.
