---
title: "Text analysis and regular expressions"
author: "Nicholas Horton (nhorton@amherst.edu)"
date: "May 19, 2022"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 5
    fig_width: 7
  word_document:
    fig_height: 3
    fig_width: 5
editor_options: 
  chunk_output_type: console
---

```{r, setup, include=FALSE}
library(tidyverse)   # Load additional packages here 
library(tidytext)
#remotes::install_github("Amherst-Statistics/DickinsonPoems")
library(DickinsonPoems)

# Some customization.  You can alter or delete as desired (if you know what you are doing).
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

Note: you'll need to:

1. install the DickinsonPoems package and
2. run `get_sentiments("afinn")` interactively

to be able to knit this.



### Warmup #1

How often does the word "science" appear in Emily Dickinson’s poetry?

```{r}
all_poems <- list_poems() %>%
  map(get_poem) %>%
  unlist() %>%
  enframe(value = "words") %>%
  unnest_tokens(word, words)
head(all_poems)
```

```{r echo = TRUE}
all_poems %>%
  filter(stringr::str_detect(word, "science"))
```

### Warmup #2

Classify Emily Dickinson's poem *The Lonely House* as either positive or negative using the `AFINN` lexicon.Does this match with your own interpretation of the poem?


```{r}
poem <- get_poem("gutenberg1.txt014")
poem
```


```{r}
# need to first run `get_sentiments("afinn")` interactively.
poem_sentiments <- data.frame(poem) %>%
  unnest_tokens(word, poem) %>%
  inner_join(get_sentiments("afinn"))

sum(poem_sentiments$value) # negative value
```


### Homework

Use the collection of Emily Dickinson poems to identify patterns or insights.
This is a very open ended assignment: you may pick a few poems or consider all of them.

